The Noisy Work of Uncertainty Visualisation Research: A Review

### 2024-04-01 paper.qmd
data-driven

There are many stages in our analysis that might benefit from the power of data visualisation, however this does not mean it is appropriately utilized or done with with effectiveness. (@Vanderplas2020 has helpful advice here.) Visualization is an important step in exploratory data analysis and it is often utilised to **learn** what is important about a data set. The importance of data driven discovery has been highlighted through the data sets Anscombe's quartet [@anscombe] and the Datasaurus Dozen [@datasaurpkg]. Each of the pairwise plots in these data sets have the same summary statistics but strikingly different information, as revealed when visualised. <!-- Anscombe quartet is shown in @fig-anscombe, because describing the data is never the same as seeing it. Instead of having to repeatedly check endless hypothesis to find interesting numerical features: DON'T NEED MUCH EMPHASIS ON THIS DATA - VERY WELL-KNOWN -->
Visualisations can **tell** us what is important about our data and **how** it might diverge from pre-conceived notions or what we expect.

@@

![*I need to fix the taxonomy illustration because inputs and model uncertainty need to be swapped*. Depicts an illustration of the taxonomy described in @utypo. From right to left the drawing shows the location, level and nature of uncertainty with examples of that category underneath. A specific source of uncertainty from the location can be mapped to a level of ignorance that can increase or decrease (i.e. moving up or down the green line) depending on the nature of the uncertainty. Identifying the location, level and nature of your uncertainty allows you to better understand it. ](taxonomyvis.jpeg){#fig-taxonomy}

>>>>

There are many stages in our analysis that might benefit from the power of data visualisation, however this does not mean it is appropriately utilized or done with with effectiveness. (@Vanderplas2020 has helpful advice here.) Visualization is an important step in exploratory data analysis and it is often utilised to **learn** what is important about a data set. The importance of data-driven discovery has been highlighted through the data sets Anscombe's quartet [@anscombe] and the Datasaurus Dozen [@datasaurpkg]. Each of the pairwise plots in these data sets have the same summary statistics but strikingly different information, as revealed when visualised. <!-- Anscombe quartet is shown in @fig-anscombe, because describing the data is never the same as seeing it. Instead of having to repeatedly check endless hypothesis to find interesting numerical features: DON'T NEED MUCH EMPHASIS ON THIS DATA - VERY WELL-KNOWN -->
Visualisations can **tell** us what is important about our data and **how** it might diverge from pre-conceived notions or what we expect.

@@

![Depicts an illustration of the taxonomy described in @utypo. From right to left the drawing shows the location, level and nature of uncertainty with examples of that category underneath. A specific source of uncertainty from the location can be mapped to a level of ignorance that can increase or decrease (i.e. moving up or down the green line) depending on the nature of the uncertainty. Identifying the location, level and nature of your uncertainty allows you to better understand it. ](taxonomyvis.jpeg){#fig-taxonomy}
